---
title: 'Examining Corresponding, Ranked Peaks'
author: 'Robert M Flight'
date: '`r Sys.time()`'
output:
  pdf_document:
    extra_dependencies: ['longtable', 'float']
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Purpose

Hunter has this idea that we can improve sample-to-sample normalization using relative ranking of peaks based on their intensity ranks.
So this is me trying to figure out if that will work.

## Setup

```{r setup_packages}
library(dplyr)
library(visualizationQualityControl)
library(smirfeTools)
library(ggplot2)
theme_set(cowplot::theme_cowplot())
```

## Load Data

Data we have is **all** the peaks from each sample, along with their ranks, along with the assigned and voted on peaks across samples.
The peaks and their IDs (`Sample_Peak`) should match up between the two data sets.

```{r load_data}
all_peaks = readRDS("testing_stuff/lung_all_peaks.rds")
voted_data = readRDS("testing_stuff/lung_voted_all.rds")
imf_data = extract_imf_emf_data(voted_data, by = "IMF")
```

## Analysis

```{r peak_correspondence}
imf_peaks = imf_data$peaks
peak_counts = apply(imf_peaks, 1, function(.x){sum(!is.na(.x))})
n_sample = ncol(imf_peaks)

peak_frac = peak_counts / n_sample
peak_frac_df = data.frame(imf = rownames(imf_peaks), fraction = peak_frac,
                          stringsAsFactors = FALSE)

imf_2_peak = purrr::map_df(rownames(imf_peaks), function(in_row){
  data.frame(imf = in_row,
             Sample_Peak = as.vector(imf_peaks[in_row, ]),
             stringsAsFactors = FALSE)
})
imf_2_peak = dplyr::filter(imf_2_peak, !is.na(Sample_Peak))
imf_2_peak = dplyr::left_join(imf_2_peak, peak_frac_df, by = "imf")
use_imf = peak_frac >= 0.9
use_peaks = imf_peaks[use_imf, ]
```

```{r peak_ranks}
n_peak_sample = dplyr::group_by(all_peaks, Sample) %>%
  dplyr::summarise(n_peak = dplyr::n())
all_peaks = dplyr::left_join(all_peaks, n_peak_sample, by = "Sample")
all_peaks$fractional_rank = 1 - (all_peaks$rank / all_peaks$n_peak)

imf_ranks = dplyr::left_join(imf_2_peak, all_peaks[, c("Sample_Peak", "fractional_rank", "n_peak")], by = "Sample_Peak")

imf_median_rank = dplyr::group_by(imf_ranks, imf) %>%
  dplyr::summarize(median_rank = median(fractional_rank))
imf_median_rank = dplyr::left_join(imf_median_rank, peak_frac_df, by = "imf")
```

Let's see if the data that Hunter would want to use even exists!

```{r check_fractions_vs_rank}
ggplot(imf_median_rank, aes(x = median_rank, y = fraction)) + geom_point(alpha = 0.5) +
  geom_vline(xintercept = 0.5, color = "red") +
  geom_hline(yintercept = 0.8, color = "red") +
  labs(subtitle = "Median Rank across Samples vs Fraction of Total Samples",
       x = "Median Rank of Peak", y = "Fraction of Samples Present")
```

OK, I previously messed up the rank within a sample, in that a lower number previously meant it was more highly ranked.
**Now it is right!**
A higher number in a sample means it is more highly ranked in the sample.

But we are going to try anyway.
The next step then is to look at the peaks that appear in a large fraction of samples (> 0.8), and compare their rank in a sample to the number of peaks in that sample.

Each of the IMFs chosen has a maximum relative rank <= 0.9 , and then we plot it's relative rank in each sample compared to the number of peaks in each sample.

```{r compare_fraction_to_rank, fig.width=8, fig.height=8}
high_fraction = dplyr::filter(imf_ranks, fraction >= 0.8)
high_fraction_maxrank = dplyr::group_by(high_fraction, imf) %>%
  dplyr::summarise(max_rank = max(fractional_rank))
has_low_rank = dplyr::filter(high_fraction_maxrank, max_rank < 0.9)

high_fraction_byrank = dplyr::left_join(has_low_rank, high_fraction, by = "imf")

p = ggplot(high_fraction_byrank, aes(x = n_peak, y = fractional_rank)) + geom_point() +
  facet_wrap(~ imf, ncol = 3, scales = "free") +
  geom_smooth()
  
p
```


```{r big_version_of_figure, include = FALSE}
Cairo::CairoPNG(filename = "testing_stuff/high_fraction_rank.png", width = 16, height = 16,
                bg = "white", units = "in", dpi = 300)
p
dev.off()
```
